---
title: "Arabic CDI - Short Forms - comprehension"
author: "Mike & George"
date: '2022-08-1'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
require(tidyverse)
require(mirt)
require(kableExtra)
library(permute)
source("IRT_helpers.R")
```

This markdown sequence documents data import of JISH and Alroqi datasets with the goals of:
1. Data wrangling
2. Short form creation
3. CAT creation

This particular document uses the combined Alroqi and JISH data to try and create an informative and useful short forms. 

Load data. 

```{r}
load(file = "cached_data/all_arabic_comprehension.Rds")
```

# Psychometric modeling

Prepare data. 

```{r}
d_comp <- select(full_comp, "sheep.sound":"also")
words <- names(d_comp)
d_mat <- as.matrix(d_comp)
```


```{r psycho-models_1pl, echo=F, eval = FALSE}
set.seed(1234)

mod_1pl <- mirt(d_mat, 1, itemtype='Rasch', verbose=TRUE, technical=list(NCYCLES=1000))

coefs_1pl <- as_tibble(coef(mod_1pl, simplify = TRUE)$items) %>%
 mutate(definition = rownames(coef(mod_1pl, simplify = TRUE)$items))
fscores_1pl <- tibble(data_id = rownames(d_mat),
 ability = fscores(mod_1pl, method = "MAP")[,1])

save(file = "cached_data/arabic_mod_1pl_comp.Rds", "mod_1pl", "fscores_1pl", "coefs_1pl")
```

```{r psycho-2pl_coefs, echo=F, eval = FALSE}
mod_2pl <- mirt(d_mat, 1, itemtype='2PL', verbose=TRUE, technical=list(NCYCLES=3000))

coefs_2pl <- as_tibble(coef(mod_2pl, simplify = TRUE)$items) %>%
 mutate(definition = rownames(coef(mod_2pl, simplify = TRUE)$items))
fscores_2pl <- tibble(data_id = rownames(d_mat),
 ability = fscores(mod_2pl, method = "MAP")[,1])

save(file = "cached_data/arabic_mod_2pl_comp.Rds", "mod_2pl","fscores_2pl", "coefs_2pl")
```


```{r psycho-fit_irt_3pl, echo=F, eval = FALSE}
mod_3pl <- mirt::mirt(d_mat, 1, itemtype='3PL', verbose=TRUE,
 technical=list(NCYCLES=4000))

coefs_3pl <- as_tibble(coef(mod_3pl, simplify = TRUE)$items) %>%
 mutate(definition = rownames(coef(mod_3pl, simplify = TRUE)$items))
fscores_3pl <- tibble(data_id = rownames(d_mat),
 ability = fscores(mod_3pl, method = "MAP")[,1])

save(file = "cached_data/arabic_mod_3pl_comp.Rds", "mod_3pl","fscores_3pl", "coefs_3pl")
```



```{r psycho-models_load, echo=F}
load("cached_data/arabic_mod_1pl_comp.Rds")
load("cached_data/arabic_mod_2pl_comp.Rds")
load("cached_data/arabic_mod_3pl_comp.Rds")
```


## Model comparison.


```{r, anovas, include=F}

mc1 <- get_anova_table(mod_1pl, mod_2pl, c("Rasch", "2PL"))
mc2 <- get_anova_table(mod_2pl, mod_3pl, c("2PL", "3PL"))
```

Compared to the Rasch model, the 2PL model fits better and is preferred by both AIC and BIC.


```{r, echo=F}
kable(mc1, digits=2,
 caption="Comparison of Rasch and 2PL models.") %>%
 html_table_width(c(60, 80, 80, 80, 50))
```

The 2PL is favored over the 3PL model.


```{r, echo=F}
kable(mc2, digits=2,
 caption="Comparison of 2PL and 3PL models.") %>%
 html_table_width(c(60, 80, 80, 80, 50))
```

We do the rest of our analyses using the 2PL model as the basis.

## Plot 2PL Coefficients
 
Next, we examine the coefficients of the 2PL model.

Items that are estimated to be very easy (e.g., mommy, daddy, ball) or very difficult (would, were, country) are highlighted, as well as those at the extremes of discrimination (a1).
 
We remove the "bad items" identified above. 

```{r}
coefs_2pl <- as_tibble(coef(mod_2pl, simplify = TRUE)$items) %>%
  mutate(definition = rownames(coef(mod_2pl, simplify = TRUE)$items)) |>
  ungroup()

ggplot(coefs_2pl,
       aes(x = a1, y = -d)) + 
  geom_point(alpha = .3) + 
  ggrepel::geom_text_repel(data = filter(coefs_2pl, 
                                a1 < 1 | a1 > 3.8 | -d > 5 | -d < -2.5), 
                  aes(label = definition), size = 2, 
                  show.legend = FALSE) + 
  xlab("Discrimination") + 
  ylab("Difficulty")
```

# Short form construction

Goal is to create a 100 item test with the best items for a given age/ability range. 

Let's find our estimated abilities and see how they relate to age. 

```{r}
# qplot(x = full_comp$age_mo, 
      # y = fscores_2pl$ability, 
      # col = full_comp$source, geom = "point" ) +
  # geom_smooth()
```

So we would like a range of abilities from about -2.5 to 4. Here's our resulting test information curve. 

```{r}
theta <- matrix(seq(-2.5,4,.01))
tinfo <- testinfo(mod_2pl, theta)
plot(theta, tinfo, type = 'l')
sum(tinfo)
```

Let's try making some random 100-item subtests. 

```{r}
coefs_2pl <- mutate(coefs_2pl, idx = 1:n())
  
tinfo_random_100 <- tibble(n = 1:1000) %>%
  split(.$n) |>
  map_df(function(x) {
    tibble(theta = as.vector(theta), 
           testinfo = testinfo(mod_2pl, theta, 
                               which.items = slice_sample(coefs_2pl, n = 100) |>
                                 pull(idx)),
           n = x$n)
  })


tinfo_random_summary <- tinfo_random_100 |>
  group_by(n) |>
  summarise(testinfo = sum(testinfo)) 

ggplot(tinfo_random_summary, 
       aes(x = testinfo)) + 
  geom_histogram()

```
The mean random test information is `r mean(tinfo_random_summary$testinfo)`.

Now let's try selecting high discrimination items. 

```{r}
top_desc <- arrange(coefs_2pl, desc(a1)) |>
  slice(1:100) |>
  pull(definition)

top_desc_idx <- which(words %in% top_desc)

tinfo_top_desc <- testinfo(mod_2pl, theta, which.items = top_desc_idx)
plot(theta, tinfo_top_desc, type = 'l')
```

The best test selecting based on discrimination has test information of `r sum(tinfo_top_desc)`. That gives us an upper bound on item information. 

Now let's try to do some kind of optimization. Our constraints are:

* We want diverse representations across categories
* We want good coverage across age

Let's look at test information for each of the categories first. 

```{r}
coefs_2pl <- left_join(coefs_2pl, 
                       items_comp |>
                         select(-definition) |>
                         rename(definition = uni_lemma) |>
                         select(definition, category))
coefs_2pl$idx <- 1:nrow(coefs_2pl)


cat_info <- tibble(cat = unique(coefs_2pl$category)) |>
  group_by(cat) |>
  mutate(data = list(tibble(theta = as.vector(theta), 
                            testinfo = testinfo(mod_2pl, theta, 
                                                which.items = filter(coefs_2pl, 
                                                                     category == cat) |>
                                                  pull(idx)),
                            n = nrow(filter(coefs_2pl, category == cat))))) |>
  unnest(cols = "data")

ggplot(cat_info, aes(x = theta, y= testinfo/n)) + 
  geom_line() + 
  facet_wrap(~cat) + 
  ggtitle("Test information per word for different sections")

```
We have 20 categories, but many of them are not so good. Let's remove the bad ones. 

```{r}
cat_info_summary <- cat_info |>
  group_by(cat) |>
  summarise(testinfo = sum(testinfo)) |>
  mutate(cat = fct_reorder(cat, testinfo)) 

cat_info_summary |>
  ggplot(aes(x = cat, y = testinfo)) +
  geom_point() + 
  coord_flip()

good_cats <- filter(cat_info_summary, testinfo > 2500) |> pull(cat)
```

Let's look at those cats.

```{r}
by_category_max_sd <- coefs_2pl |>
  ungroup() |>
  filter(category %in% good_cats) %>%
  split(.$category) |>
  map_df(function (cat) {
    
    perms <- tibble(n = 1:100) %>%
      split(.$n) |>
      map_df(function (x) {
        slice_sample(cat, n=5) |>
          mutate(score = sd(d) + mean(a1),
                 n = x$n)
      })
    
    filter(perms, score == max(score))
  })
  
by_cat_max_sd_test_info <- testinfo(mod_2pl, theta, which.items = by_category_max_sd$idx)
```

Let's try just maximizing discrimination within category. 

```{r}
by_category_max_desc <- coefs_2pl |>
  ungroup() |>
  filter(category != "Sound Effects and Animal Sounds") %>%
  split(.$category) |>
  map_df(function (cat) {
    
    arrange(cat, desc(a1)) |>
      slice(1:5)
  })

by_cat_max_desc_test_info <- testinfo(mod_2pl, theta, which.items = by_category_max_desc$idx)
```

How about adding the easiest one in each category and then doing the four most discriminating.

```{r}
by_category_max_desc_one_easy <- coefs_2pl |>
  ungroup() |>
  filter(category != "Sound Effects and Animal Sounds") %>%
  split(.$category) |>
  map_df(function (cat) {
    
    # do this so we definitely get 5 even if the two conditions overlap
    filter(cat, d==max(d) | a1 >= sort(a1, decreasing=TRUE)[5]) |>
      arrange(desc(d)) |>
      slice(1:5)
      
  })

by_category_max_desc_one_easy_test_info <- testinfo(mod_2pl, theta, 
                                                    which.items = by_category_max_desc_one_easy$idx)
```


Now compare these. 

```{r}
sf_vs_best <- tibble(theta = theta, 
                     `balance discrim and difficulty by category` = by_cat_max_sd_test_info, 
                     `one easy plus most discrim by category` = by_category_max_desc_one_easy_test_info,
                     `most discrim by category` = by_cat_max_desc_test_info, 
                     `most discrim overall` = tinfo_top_desc, 
                     `random` = tinfo_random_100 |> 
                       group_by(theta) |> 
                       summarise(testinfo = mean(testinfo)) |> 
                       pull(testinfo)) |>
  pivot_longer(-theta, names_to = "selection", values_to = "test_information")
                       
                       
ggplot(sf_vs_best, 
       aes(x = theta, 
           y = test_information, col = selection)) + 
  geom_line() 
  
```

Let's adopt the most discrim by category option here since it is generally better...

# Examine resulting items

```{r}
short_metrics <- full_comp |>
  rowwise() |>
  mutate(top_desc = sum(c_across(cols = coefs_2pl$definition[top_desc_idx])),
    by_cat = sum(c_across(cols = 
                              coefs_2pl$definition[by_category_max_desc$idx]))) |>
  select(subid, age_mo, total, top_desc, by_cat)

short_metrics_long <- short_metrics |>
  pivot_longer(top_desc:by_cat, names_to = "selection", values_to = "estimate")
  

ggplot(short_metrics_long, aes(x = total, y = estimate)) +
  geom_point(alpha = .25) +
  geom_smooth() + 
  geom_abline(lty = 2, slope = 100/length(coefs_2pl$definition)) +
  facet_wrap(~selection)
```

Correlations. 

```{r}
cor.test(short_metrics$total, short_metrics$top_desc)
with(filter(short_metrics, age_mo <= 18), 
     cor.test(total, top_desc))

```
Correlations are very high for the whole sample, but lower for younger kids. 

```{r}
cor.test(short_metrics$total, short_metrics$by_cat)
with(filter(short_metrics, age_mo <= 18), 
     cor.test(total, by_cat))
```
Our corrected test ("by cat") does substantially better with the younger kids. 

What's in that test? 

```{r}
filter(coefs_2pl, idx %in% by_category_max_desc_one_easy$idx) |>
  select(definition, category, d, a1) |>
  DT::datatable()
```

We can maybe do better by removing some redundancy (e.g. toy/toys), but this doesn't look totally crazy. 

# Output items

```{r}
item_params <- left_join(coefs_2pl, 
                         items_comp |> 
                           select(definition, uni_lemma) |>
                           rename(arabic = definition, 
                                  definition = uni_lemma)) |>
  rename(difficulty = d, 
         discrimination = a1) |>
  select(idx, category, arabic, definition, difficulty, discrimination)

write_csv(item_params, "full_word_list_with_parameters_comp.csv")
write_csv(filter(item_params, 
                 idx %in% by_category_max_desc$idx), 
          "candidate_100_item_production_form_list_comps.csv")
```

