---
title: "Percentiles"
author: "Arabic CDI team"
date: "2023-4-27"
output: 
  html_document:
    code_folding: hide
---

The goal of this markdown is to show demographic data analyses and percentile tables for the YNMO short form CDI data. 

# Preliminaries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
library(gamlss)
require(tidyverse)
require(mirt)
require(readxl)
library(here)
select <- dplyr::select
```

Read in data. 

```{r}
ynmo1 <- suppressMessages(read_xlsx(path="raw_data/YNMO/Ynmo Norms Part 1 Sept 2022 - Final.xlsx", 
                                       sheet="Part 1 responses")) |>
  mutate(part = 1) |>
  slice(-1)

ynmo2 <- suppressMessages(read_xlsx(path="raw_data/YNMO/Ynmo Norms Part 2 Jan 2023 - Final .xlsx", 
                                       sheet=" Part 2 responses")) |>
  mutate(part = 2) |>
  slice(-1)

ynmo3 <- suppressMessages(read_xlsx(path="raw_data/YNMO/Ynmo Norms Part 3 Jan 2023 -Final.xlsx", 
                                       sheet="Part 3 Responses - Ynmo app")) |>
  mutate(part = 3) |>
  slice(-1)

raw_ynmo <- bind_rows(ynmo1, ynmo2, ynmo3)
```

Consider ages in the removed data. 

```{r}
qplot(raw_ynmo$age_in_months, fill = as.factor(raw_ynmo$part), binwidth = 6)
```



```{r}
ynmo <- raw_ynmo

demo <- ynmo |>
  select(subject_ID : disability_or_language_disorders_other_family)

ynmo_wg <- ynmo |>
  filter(!(`cat...117` %in% c("yes","no"))) |>
  mutate(form = "WG")

ynmo_ws <- ynmo |>
  filter(`cat...117` %in% c("yes","no")) |>
  mutate(form = "WS")
```

Overall totals.

```{r}
bind_rows(ynmo_wg, ynmo_ws) |>
  group_by(form) |>
  count() |>
  knitr::kable()
```

By part and form totals.

```{r}
bind_rows(ynmo_wg, ynmo_ws) |>
  group_by(form, part) |>
  count() |>
  knitr::kable()
```



# Words and Gestures

There is not enough data on WG to move froward yet. 

```{r}
ynmo_wg_long <- ynmo_wg |>
  select(subject_ID, age_in_months, sex, `cat...25`:`also...116`) |>
  pivot_longer( `cat...25`:`also...116`, names_to = "item", values_to = "value") |>
  mutate(produces = ifelse(value %in% "produces", 1, 0), 
         understands = ifelse(value %in% c("understands","produces"), 1, 0))


ynmo_wg_ms <- ynmo_wg_long |>
  group_by(subject_ID, age_in_months, sex) |>
  summarise(understands = sum(understands), 
            produces = sum(produces))
```

## Production

92 items on wg, 100 on WS

```{r}
max_vocab = 92
quantiles = c(10, 25, 50, 75, 90)

ynmo_wg_ms$understands_01 <- .001 + ynmo_wg_ms$understands/(max_vocab+1) 
ynmo_wg_ms$produces_01 <- .001 + ynmo_wg_ms$produces/(max_vocab+1) 

prod_mod_data <- filter(ynmo_wg_ms, age_in_months >= 8, age_in_months < 24)

ynmo_wg_prod_mod <- gamlss(produces_01 ~ pbm(age_in_months, lambda = 10000),
              sigma.formula = ~ pbm(age_in_months, lambda = 10000),
              family = BE, 
              control = gamlss.control(c.crit = .1),
              data = prod_mod_data)

centiles <- centiles.pred(ynmo_wg_prod_mod, cent = quantiles,
              xname = "age_in_months", xvalues = 8:18) |>
  rename(age_in_months = x) |>
  pivot_longer(-age_in_months, names_to = "percentile", values_to = "predicted") |>
  mutate(predicted = (predicted * (max_vocab+1)) - .001) 
```


```{r}
ggplot(prod_mod_data, aes(x = age_in_months, y = produces)) + 
  geom_jitter(width = .2, height = 0, alpha = .2) + 
  geom_line(data = centiles, aes(x = age_in_months, y = predicted, col = percentile)) +
  xlim(8,18)
```


## Comprehension

```{r}
comp_mod_data <- filter(ynmo_wg_ms, age_in_months >= 8, age_in_months < 24)

ynmo_wg_comp_mod <- gamlss(understands_01 ~ pbm(age_in_months, lambda = 10000),
              sigma.formula = ~ pbm(age_in_months, lambda = 10000),
              family = BE, 
              control = gamlss.control(c.crit = .1),
              data = comp_mod_data)

centiles <- centiles.pred(ynmo_wg_comp_mod, cent = quantiles,
              xname = "age_in_months", xvalues = 8:18) |>
  rename(age_in_months = x) |>
  pivot_longer(-age_in_months, names_to = "percentile", values_to = "predicted") |>
  mutate(predicted = (predicted * (max_vocab+1)) - .001) 

ggplot(ynmo_wg_ms, aes(x = age_in_months, y = understands)) + 
  geom_jitter(width = .2, height = 0, alpha = .2) + 
  geom_line(data = centiles, aes(x = age_in_months, y = predicted, col = percentile)) + xlim(8,18)
```

# Words and Sentences

```{r}
ynmo_ws_long <- ynmo_ws |>
  select(subject_ID, age_in_months, sex, `cat...117`:`if.2`, combine_words) |>
  pivot_longer( `cat...117`:`if.2`, names_to = "item", values_to = "value") |>
  mutate(produces = ifelse(value %in% "yes", 1, 0))


ynmo_ws_ms <- ynmo_ws_long |>
  group_by(subject_ID, age_in_months, sex) |>
  summarise(produces = sum(produces), combine_words = combine_words[1])

ggplot(filter(ynmo_ws_ms, age_in_months >= 16, age_in_months <= 36), 
       aes(x = age_in_months, y = produces)) + 
  geom_jitter(width = .2, height = 0, alpha = .2) + 
  geom_smooth() 
```

Consider sex effects.

```{r}
ggplot(filter(ynmo_ws_ms, age_in_months >= 16, age_in_months <= 36), 
       aes(x = age_in_months, y = produces, col = sex)) + 
  geom_jitter(width = .2, height = 0, alpha = .2) + 
  geom_smooth() 
```

Now with percentiles. 

```{r}
max_vocab = 100
quantiles = c(10, 25, 50, 75, 90)

ynmo_ws_ms$produces_01 <- .001 + ynmo_ws_ms$produces/(max_vocab+1) 


ws_mod_data <- filter(ynmo_ws_ms, age_in_months >= 16, age_in_months <= 36)

ynmo_ws_prod_mod <- gamlss(produces_01 ~ pbm(age_in_months, lambda = 10000),
              sigma.formula = ~ pbm(age_in_months, lambda = 10000),
              family = BE, 
              control = gamlss.control(c.crit = .1),
              data = ws_mod_data)

centiles <- centiles.pred(ynmo_ws_prod_mod, cent = quantiles,
              xname = "age_in_months", xvalues = 16:36) |>
  rename(age_in_months = x) |>
  pivot_longer(-age_in_months, names_to = "percentile", values_to = "predicted") |>
  mutate(predicted = (predicted * (max_vocab+1)) - .001) 


ggplot(ynmo_ws_ms, aes(x = age_in_months, y = produces)) + 
  geom_jitter(width = .2, height = 0, alpha = .2) + 
  geom_line(data = centiles, aes(x = age_in_months, y = predicted, col = percentile)) + 
  xlim(16,36) +
  ylab("Total productive vocabulary") + 
  xlab("Age (months)") +
  scale_color_discrete(name = "Percentile") + 
  ggthemes::theme_few()

ggsave(here("graphs","percentiles.png"), width = 6, height = 4)
```


```{r}

ynmo_ws_prod_sex_mod <- gamlss(produces_01 ~ pbm(age_in_months, lambda = 10000) * sex,
              sigma.formula = ~ pbm(age_in_months, lambda = 10000),
              family = BE, 
              control = gamlss.control(c.crit = .1),
              data = ws_mod_data)

pred_data <- expand_grid(age_in_months = 16:36, 
                         sex = c("male","female"))
pred_data$pred <- predict(ynmo_ws_prod_sex_mod, 
                          type = "response",
                          newdata = pred_data) 

pred_data <- pred_data |>
  mutate(vocab_pred = (pred * (max_vocab+1)) - .001, 
         sex = ifelse(sex == "female","Female","Male")) 


ggplot(ynmo_ws_ms |>
         mutate(sex = ifelse(sex == "female","Female","Male")),
       aes(x = age_in_months, y = produces, col = sex)) + 
  geom_jitter(width = .2, height = 0, alpha = .2) + 
  geom_line(data = pred_data, aes(x = age_in_months, 
                                  y = vocab_pred, col = sex)) + 
  xlim(16,36) +
  ylab("Total productive vocabulary") + 
  xlab("Age (months)") +
  ggthemes::scale_color_solarized(name = "Sex") + 
  ggthemes::theme_few()

ggsave(here("graphs","sex.png"), width = 6, height = 4)
```


## Combining words

```{r}
combines_data <- ws_mod_data |>
  mutate(`not yet` = ifelse(combine_words == "not yet", 1, 0),
         sometimes = ifelse(combine_words == "sometimes", 1, 0), 
         often = ifelse(combine_words == "often", 1, 0)) |>
  pivot_longer(`not yet`:often, names_to = "combines_words", 
               values_to = "value") |>
  mutate(combines_words = factor(combines_words, levels = c("not yet","sometimes","often")))

ggplot(combines_data, 
       aes(x = age_in_months, y = value, col = combines_words)) + 
  geom_smooth(method = "glm", formula = y ~ x, method.args = list(family = "binomial")) + 
  ylab("Proportion of children") + 
  xlab("Age (months)") +
  ggthemes::scale_color_solarized(name = "Combines words") + 
  ggthemes::theme_few()

ggsave(here("graphs","combines.png"), width = 6, height = 4)

```




```{r}
ynmo_ws_prod_combine_mod <- gamlss(produces_01 ~ pbm(age_in_months, lambda = 10000) + combine_words,
              sigma.formula = ~ pbm(age_in_months, lambda = 10000),
              family = BE, 
              control = gamlss.control(c.crit = .1),
              data = ws_mod_data)

pred_data <- expand_grid(age_in_months = 16:36, 
                         combine_words = c("not yet","sometimes","often"))
pred_data$pred <- predict(ynmo_ws_prod_combine_mod, 
                          type = "response",
                          newdata = pred_data) 

pred_data <- pred_data |>
  mutate(vocab_pred = (pred * (max_vocab+1)) - .001) |>
  mutate(combine_words = factor(combine_words, 
  levels = c("not yet","sometimes","often")))


ggplot(ynmo_ws_ms |>
  mutate(combine_words = factor(combine_words, 
                                levels = c("not yet",
                                           "sometimes","often"))), 
  aes(x = age_in_months, y = produces, 
                       col = combine_words)) + 
  geom_jitter(width = .2, height = 0, alpha = .2) + 
  geom_line(data = pred_data, aes(x = age_in_months, 
                                  y = vocab_pred, col = combine_words)) + 
  xlim(16,36) +
  ylab("Total productive vocabulary") + 
  xlab("Age (months)") +
  ggthemes::scale_color_solarized(name = "Combines words") + 
  ggthemes::theme_few()

ggsave(here("graphs","combine_vocab.png"), width = 6, height = 4)
```

```{r}
summary(ynmo_ws_prod_combine_mod)
```


Combines table.

```{r}
ws_mod_data |>
  mutate(`not yet` = ifelse(combine_words == "not yet", 1, 0),
         sometimes = ifelse(combine_words == "sometimes", 1, 0), 
         often = ifelse(combine_words == "often", 1, 0)) |>
  group_by(age_in_months, sex) |>
  summarise(`not yet` = mean(`not yet`), 
            `sometimes or often` = mean(sometimes | often)) |>
  pivot_wider(names_from = "sex", values_from = c("not yet","sometimes or often")) |>
  write_csv("outputs/WS_combines.csv")

```



# other analysis

Compute MADM. 

```{r}
ynmo_ws_ms |>
  mutate(age_binned = floor(age_in_months / 2)*2) |>
  group_by(age_binned) |>
  summarise(madm = mean(abs(produces - median(produces))) / median(produces)) |>
  ggplot(aes(x = age_binned, y = madm)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```

Output percentile tables

```{r}
all_quants = seq(5,95,5)

cents <- centiles.pred(ynmo_ws_prod_mod, cent = all_quants,
              xname = "age_in_months", xvalues = 16:36) |>
  rename(age_in_months = x) |>
  pivot_longer(-age_in_months, names_to = "percentile", values_to = "produces") |>
  mutate(produces = round((produces * (max_vocab+1)) - .001), 
         produces = ifelse(produces > 100, 100, produces))  |>
  pivot_wider(names_from = "percentile", values_from = "produces")

write_csv(cents, "wg_centiles_2023-05-04.csv")
```

Output percentile tables

```{r}

ws_mod_data_m <- filter(ws_mod_data, sex == "male")
ynmo_ws_prod_mod_m <- gamlss(produces_01 ~ pbm(age_in_months, lambda = 10000),
                               sigma.formula = ~ pbm(age_in_months, lambda = 10000),
                               family = BE, 
                               control = gamlss.control(c.crit = .1),
                               data = ws_mod_data_m)

ws_mod_data_f <- filter(ws_mod_data, sex == "female")
ynmo_ws_prod_mod_f <- gamlss(produces_01 ~ pbm(age_in_months, lambda = 10000),
                               sigma.formula = ~ pbm(age_in_months, lambda = 10000),
                               family = BE, 
                               control = gamlss.control(c.crit = .1),
                               data = ws_mod_data_f)

cents_m <- centiles.pred(ynmo_ws_prod_mod_m, cent = all_quants,
                       xname = "age_in_months", xvalues = 16:36) |>
  rename(age_in_months = x) |>
  pivot_longer(-age_in_months, names_to = "percentile", values_to = "produces") |>
  mutate(produces = round((produces * (max_vocab+1)) - .001), 
         produces = ifelse(produces > 100, 100, produces))  |>
  pivot_wider(names_from = "percentile", values_from = "produces")

cents_f <- centiles.pred(ynmo_ws_prod_mod_f, cent = all_quants,
                       xname = "age_in_months", xvalues = 16:36) |>
  rename(age_in_months = x) |>
  pivot_longer(-age_in_months, names_to = "percentile", values_to = "produces") |>
  mutate(produces = round((produces * (max_vocab+1)) - .001), 
         produces = ifelse(produces > 100, 100, produces))  |>
  pivot_wider(names_from = "percentile", values_from = "produces")

write_csv(cents_m, "wg_centiles_male_2023-05-04.csv")
write_csv(cents_f, "wg_centiles_female_2023-05-04.csv")
```
